
# Introduction

```{r}
#| message: false
library(tidyverse)
library(gt)
```

## Summary

Accurate estimates of size at maturity are crucial for the effective management of marine crustaceans, which represent the most valuable group of wild-capture fisheries by landed mass. Size at maturity values often influence the establishment of minimum landing sizes and represent crucial inputs in widely-used stock assessment models. For many crustaceans, size at maturity can be observed through changes in the relative size and growth of the chelae (males) or abdomen (females), referred to as a transition to morphometric maturity. Morphometric maturity is often a better indicator of functional maturity (i.e., the ability to successfully mate and produce offspring) than physiological maturity, which generally must be determined through gonad dissection or similarly invasive methods. Over the past century, various methods have been developed to estimate size at morphometric maturity for hundreds of crustacean species across the globe. Despite their widespread use, there has been limited research comparing the accuracy of these methods, and many researchers are unaware of more recently developed approaches to modeling size at maturity using morphometric data. My proposed thesis will address these gaps by comprehensively evaluating all documented approaches to modeling size at morphometric maturity and developing tools to facilitate the adoption of sound quantitative methods by fisheries scientists and managers.

## Proposed Structure

I propose that my dissertation will contain at least five chapters, organized under three main headings: synthesis research (literature reviews and meta-analyses), quantitative methods (simulation testing and R package development), and applications to specific crustacean species (@tbl-outline).

```{r}
#| label: tbl-outline
#| tbl-cap: "Proposed structure of PhD dissertation"

tbl_outline <- tribble(
  ~"Heading", ~"Chapter", ~"Title",
  "Synthesis", "Ch. 1", "A systematic review of the use of morphometric data to determine crustacean size at maturity",
  "Synthesis", "Ch. 2", "Ecological and evolutionary drivers of changes in crab (Brachyura) allometry at maturity",
  "Quantitative methods", "Ch. 3", "A guide to determining crustacean size at maturity from morphometric data (simulation testing)",
  "Quantitative methods", "Ch. 4", "morphmat: An R package to Model Size at Maturity Using Morphometric Data",
  "Applications", "Ch. 5", "Modeling spatial variation in Jonah crab size at maturity",
  "Applications", "Ch. 6", "TBD additional collaborations/applications. May be research output from PhD but not formally included in dissertation"
)

tbl_outline %>% gt() %>% tab_options(column_labels.hidden = TRUE)
```


My proposed chapters are thematically connected and build upon one another in a natural progression. Chapter 1 will focus on identifying the modeling approaches (e.g., various types of piecewise regression and clustering) used to estimate size at maturity and will examine the context in which they are applied and the conclusions drawn from the given application. Chapter 2 will use the comprehensive dataset compiled in Chapter 1 to identify how changes in morphometry may be correlated with the life history traits, ecological niches, and evolutionary trajectories of different crustacean species. Chapter 3 will use simulation testing to evaluate the performance of the models identified in Chapter 1 when applied to artificial datasets with known size at maturity values. The simulation testing will allow me to determine how accurately each model can recover the true parameters when presented with realistic challenges like small sample sizes, underrepresentation of smaller individuals, and high variability/error in the data.

Chapter 4 describes an open-source R package I am developing that will contain easy-to-implement functions for researchers to implement the modeling approaches identified and tested in the previous chapters. Finally, the Applications section will demonstrate the value of this research by using the best practices I identify in Chapter 3 to model size at maturity for specific crustacean species. At a minimum, this will include a chapter modeling spatial variation in Jonah crab (*Cancer borealis*) size at maturity, illustrating how the workflows discussed under the Quantitative Methods heading can be extended to incorporate spatially-structured or time series data. However, I have identified numerous potential avenues to apply this work in collaboration with external researchers (@tbl-possiblecollabs), and undoubtedly more will arise over the course of my graduate studies.

```{r}
#| label: tbl-possiblecollabs
#| tbl-cap: "Example extensions/applications of the topics discussed in the main chapters."

collabs <- tribble(
  ~ "Topic",  ~ "Potential collaborators",
  "Identifying regenerating chelae when analyzing crustacean morphometrics",
  "Jason Goldstein (Wells NERR, UNH), Joshua Carloni (NH F&G)",
  "Climate impacts on size at morphometric maturity for Atlantic crustacean fisheries",
  "Kathy Mills, Michelle Staudinger, Bernard Sainte-Marie",
  "Informing Jonah crab and American lobster population models and stock assessments",
  "Miguel Barajas, Jesica Waller, Everett Rzeszowski, Andrew Goode, Cameron Hodgdon",
  "Morphometric vs. physiological maturity of the Chesapeake Bay blue crab",
  "Zach Olsen (Texas PWD), VIMS/UMCES (Alexandra Schneider)"
)

collabs %>% gt()
```


## Morphometric Size at Maturity - Background

### Types of Maturity

### Types of Allometric Growth in Crustaceans

## References {.unnumbered}
