---
bibliography: references.bib
---

# Systematic Review

```{r}
#| message: false
library(tidyverse)
library(gt)
```

## Introduction

The foundation of my dissertation will be a systematic literature review to identify all modeling approaches used to estimate crustacean size at maturity using morphometric data. Special focus will be placed on understanding the methodologies of studies that have investigated regional/spatial variation in size at maturity. I will aim to follow current best practices for ensuring transparency and reproducibility in ecological systematic reviews and meta-analyses [@foo2021; @odea2021; @kambouris2024; @gates2002; @koricheva2013].

Synthesis research plays a crucial role in accelerating ecological knowledge by integrating evidence and organizing information in a way that can answer foundational, large-scale questions [@halpern2020; @spake2022]. The combined analysis of disparate data sources, theories, and tools is a powerful tool for advancing scientific understanding, and represents a powerful approach to informing evidence-based management decisions in the face of rapid environmental change [@halpern2020; @koricheva2013]. The results from qualitative and, where possible, quantitative/meta-analytical methods applied to a novel database of size at maturity studies will be invaluable for preventing the remainder of my graduate work from duplicating existing efforts and maximizing the scientific value of my research output. Additionally, given the relative impact of synthesis papers compared to non-synthesis papers, publishing a rigorous systematic review will be highly beneficial for establishing my presence as a promising researcher within the fisheries science community.

*Section about importance of reproducible synthesis research & how this project aligns with the NMFS Openscapes initiative and general movement towards open science*

## Completed work

### Formulate initial question and identify inclusion/exclusion criteria

*Overarching question:* What methods have previously been used to model crustacean size at maturity based on morphometric data, and in what contexts have they been applied?

Since the focus of this review does not fit well into the PICO framework, I have used the SPIDER framework to organize the key components of the research question [@methley2014] (@tbl-spider).

```{r}
#| label: tbl-spider
#| tbl-cap: "Inclusion criteria in the SPIDER framework."

spider <- tribble(
  ~ letter,  ~ word,         ~ description,
  "S",       "Sample",       "Crustacean species",
  "PI",      "Phenomenon of Interest",
  "Identifying size at sexual maturity based on changes in a morphometric relationship (e.g., the ratio between carapace width and claw height)",
  "D" ,      "Design",
  "Measured at least two morphological features, then used a computational/algorithmic approach to estimate size at morphometric maturity",
  "E",       "Evaluation",
  "A numerical estimate of SM50 or conclusion about the utility of morphometric analysis for determining size at maturity",
  "R",       "Research Type", "Quantitative"
)

spider %>% gt() %>% tab_options(column_labels.hidden = TRUE)
```

[Exclusion criteria:]{.underline}

-   Studies that did not attempt to estimate size at maturity or identify maturity-related changes in allometric growth
    -   Studies that used existing size cutoffs (e.g., 30mm CW) or discriminant functions for classification
-   Studies that only estimate size at gonadal/sexual maturity (no morphometric ratios considered)
-   Studies that focused on a non-crustacean species
-   Studies written in a language other than English

### Execute search

To identify the most appropriate citation databases to use for this review, I consulted several studies comparing the academic search systems commonly used for evidence synthesis [@visser2021; @haddaway2015; @martín-martín2021; @gusenbauer2019; @gusenbauer2020]. Based on the overlap between the databases recommended for use by these studies and the databases to which I have institutional access, I decided to use Scopus, JSTOR, and ProQuest. Although the merits of Google Scholar as a tool for systematic reviews may be somewhat contentious [@boeker2013; @giustini2013], it contains many records that are not found in other databases and is a powerful tool to find grey literature and increase the coverage of multi-database searches [@haddaway2015; @gusenbauer2019; @martín-martín2021]. I only included the first 500 results out of 686 returned from Google Scholar (the first 25 pages). This is beyond the threshold of 50-100 results included by many systematic reviewers [@haddaway2015], and results beyond that appeared to have low relevance to the focal question. I also searched the NOAA Institutional Repository, another excellent source of grey literature, and did preliminary searches of Github and Zenodo to identify code/software developed for size at maturity modeling. Additionally, many additional relevant sources were identified through snowball searching and while reading scientific articles and management documents (e.g., stock assessment reports) (@tbl-databases).

```{r}
#| label: tbl-databases
#| tbl-cap: "Databases used for systematic review"

databases <- tribble(
  ~Database,     ~"# of results", ~"Search field",             ~Notes,
  "JSTOR",        33,          "Abstracts only",            "",
  "Scopus",       332,         "Abstract, title, keywords", "",
  "ProQuest",     49,           "Anywhere but full text",   "ProQuest One Academic (includes ProQuest Central and ProQuest Dissertation & Theses Global. Chose source types: Conference Papers & Proceedings, Dissertations & Theses, Books, Encyclopedias & Reference Works, Reports, Scholarly Journals, Trade Journals, and Working Papers",
  "Google Scholar",500,           "Title",          "Excluded results that were available only as citations",
  "NOAA Institutional Repository",392,"Everything",               "",
  "Snowball & other",139,           NA,              "Sources found in the reference list (backward snowballing) or citations (forward snowballing) of screened papers, or those identified in the process of doing background research or general reading for the project."
)

databases %>% 
  gt()
```

Developing a formal search strategy is a key component of a systematic review [@foo2021]. For Google Scholar, the search query used was: intitle:(("maturity" AND "size") AND ("crab" OR "Crustacea" OR "Decapoda" OR "Brachyura")). The query used for all other databases was: (morpho\* AND "maturity" AND "size") AND ("crab" OR "Crustacea" OR "Decapoda" OR "Brachyura"). The search string was tested with a preliminary Scopus search to ensure that it could find two articles that were highly relevant to the project: Olsen and Stevens [-@olsen2020] and Lawrence et al. [-@lawrence2021].

Duplicates were first removed using the duplicate merging feature in the reference manager software Zotero. A second round of duplicate screening was conducted using the online systematic review software Rayyan, which was then used for the initial screening of titles and abstracts. In some cases, the full text was briefly skimmed to resolve uncertainties about whether the study should be included that resulted from absent or ambiguous descriptions in the abstract of the method of determining size at maturity. The primary reason for exclusion was that the study determined physiological or functional maturity rather than morphometric maturity [@waiho2017]. Another common category of excluded studies were those using morphometric analysis for a purpose other than determining size at maturity, such as to distinguish two species, compare populations within a species, or identify the biological effects of an external variable like temperature or predation. At least 70 studies were focused on the general morphology and/or size at maturity of a non-crustacean organism, particularly skates, rays, and marine snails.

A small number of studies were excluded because they were written in a language other than English. Although not considering non-English studies can introduce bias into systematic reviews, some authors have found excluding non-English papers has a minimal effect on the overall conclusions of evidence syntheses [@vickers1998; @nussbaumer-streit2020; @jüni2002]. Additionally, one of the major arguments against excluding non-English studies is that it skews the geographical provenance of the body of evidence under consideration towards Europe and North America [@walpole2019]. The \>500 English studies retained for future screening represented a wide geographic distribution, and nearly all non-English studies were conducted in a region already represented by multiple English studies.

## Data extraction

Building a well-designed database is a fundamental component of a comparative analysis of ecological data and allows the data to be reused and expanded in support of future scientific inquiry [@schwanz2022].

My database will include three main data tables for extracted data alongside three supporting tables with important metadata.

One will be in "wide" format with one row for each paper/source. The Parameters and Species tables will be in long/stacked format, with one row for each relevant grouping included in a source.

Following EML format, the first supporting table contains factor codes. The second supporting table contains standardized descriptions of each measured body part to mitigate confusion resulting from the usage of general terms like "claw height". Similarly, a third table contains standardized descriptions of each modeling approach to mitigate confusion resulting from the usage of general terms like "piecewise regression".

Additional information regarding the storage and accessibility of these data is available in @sec-datamgmt.

## Publication bias

In any systematic review, it is necessary to investigate potential bias in the studies included in the review. I will use appropriate graphical and statistical tests (e.g., funnel plots) to identify the existence of publication bias and assess the impact of publication bias on the results of the review [@nakagawa2012; @nakagawa2022]. I will also discuss potential bias associated with the literature search and extraction proceses [@koricheva2013].

## Data analysis

Most factor variables will be summarized by calculating the proportion of included studies for which the variable was coded TRUE: for example, if the study also recorded physiological maturity, if the study included a measure of variance for their SM50 estimate, or if the study mentioned excluding individuals with regenerating chelae.

More complex meta-analytical models will be fit as Bayesian multilevel mixed models, likely using the `MCMCglmm` or `brms` packages with `bayestestR` or `bayesplot` for visualizing posterior distributions, trace plots, etc. These models will include phylogenetic correlation matrices and can account for non-independence when multiple effect sizes are obtained from the same study.

## References {.unnumbered}
